Efficient GPU implementation of multicomponent and multiphase Lattice Boltzmann models.

### 2010-04-05 paper.tex
Start a section about the CUDA programming model.

\section{The Lattice-Boltzmann method}
\label{sec:lbm:binary:liquids}
This works concentrates on two popular LBM multiphase models. Originally introduced
in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers
by its simplicity.  The free-energy model \cite{swift} is based on the free-energy
phase transition \cite{landau}. Both models werre successfully applied in simulation
of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels
\cite{pooley-contact}, thermal flow \cite{zhang-thermal}, and studies of break-up
of the liquid droplet \cite{nourgaliev-breakup}. The section describes the
two-dimensional implementation of the Shan-Chen and free-energy models with
the D2Q9 model.

@@

\section{GPU implementation}
\section{Results}
\label{sec:benchmark}

>>>>

\section{The Lattice-Boltzmann method}

{\color{red} a general overview of basics of LBM needs to be added here (grid, collision, propagation, etc)}

\label{sec:lbm:binary:liquids}
This works concentrates on two popular LBM multiphase models. Originally introduced
in 1993 \cite{Shan-chen:extended} the model attracted a multitude of researchers
by its simplicity.  The free-energy model \cite{swift} is based on the free-energy
phase transition \cite{landau}. Both models werre successfully applied in simulation
of two-phase blood flow \cite{halliday-multicomponent}, flow in microchannels
\cite{pooley-contact}, thermal flow \cite{zhang-thermal}, and studies of break-up
of the liquid droplet \cite{nourgaliev-breakup}. The section describes the
two-dimensional implementation of the Shan-Chen and free-energy models with
the D2Q9 model.

@@

\section{The CUDA Programming Environment}

CUDA is implemented in the video card hardware by organizing the GPU around the
concept of a multiprocessor (MP).  Such a multiprocessor consists of several (8 for
devices of compute capability lower than 2.0) scalar processors (SPs), each of which is
capable of executing a thread in a SIMT (Single Instruction, Multiple Threads) manner.
Each MP also has a limited amount of specialized on-chip memory: a set of 32-bit registers,
a shared memory block, a constant cache and a texture cache.  The registers are local
to the scalar processor, but the other types of memory are shared between all SPs
in a MP, which allows data sharing between threads.

{\color{red} TODO: verify the figures in this paragraph}
Perhaps the most salient feature of the CUDA architecture is the memory hierarchy
with 1-2 orders of magnitude differences between access times at each successive levels.

The slowest kind of memory is the host memory accessible only to the CPU.  This memory,
while potentially very large in size, cannot be accessed directly by the GPU, so in order
for data stored there to be useful during a GPU simulation, it has to be copied to the
global device memory.  Such a data transfer goes over the PCI-e bus, which has a maximum
throughput of about 6~GB/s.

Next in line is the global device memory of the GPU, which can currently span several
gigabytes and which has a bandwidth of about 100~GB/s.  Global memory accesses are however
high-latency operations, taking several hundred clock cycles of the GPU to complete {\color{red} Cite NVIDIA documentation here}

The fastest kind of memory currently available on GPUs is the shared memory block residing
on MPs.  It is currently limited in size to just 16~kB, but has a bandwidth of ca~1.3~TB/s
and usually a latency no higher than that of a~SP register access.

The above description readily suggests an optimization strategy which we will generally
follow in the next section and which can be summarized as: move as much data as possible
to the fastest kind of memory available and keep it there as long as possible, while minimizing
accesses to the slower kinds of memory.

\section{GPU implementation of LBM}

We will now outline an efficient and general GPU LBM algorithm taking into consideration
the facts from the previous section.

\section{Results}
\label{sec:benchmark}
